---
title: "Penguin Size Comparison"
author: "Iris Foxfoot"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) + geom_histogram(bins = 10) + facet_wrap(~sex)

#Quantile Quantile Plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) + geom_qq() + facet_wrap(~sex)

#Descriptive Statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarise(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n())
```

Get vectors of body_mass_g for female and male chinstrap penguins separately. 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

## T-tests, a couple of different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```
Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
``` 

## One-way ANOVA

## 1. Table of counts by penguin species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```
Create a subset of only female penguins
```{r}
penguins_f <- penguins %>%
  filter(sex == "female")

ggplot(data = penguins_f, aes(x = flipper_length_mm)) + geom_histogram(aes(fill = species)) + facet_wrap(~species, scales = "free")

ggplot(data = penguins_f, aes(sample = flipper_length_mm)) + geom_qq(aes(color = species)) + facet_wrap(~species, scales = "free")
```
 ### Find Summary Statistics
 
 starting from penguins_f, create a summary table that groups by speceies and returns the following:
 
 mean flipper length
 
 sd""
 
 sample size ""
 
 variance 'var' 
 
 standard error = sd/sqrt(n)
 
```{r}
penguin_f_summary <- penguins_f %>%
  group_by(species) %>% 
  summarise(mean_fl = mean(flipper_length_mm), sd_fl = sd(flipper_length_mm), sample_size = n(), varience_fl = var(flipper_length_mm), se_fl = sd(flipper_length_mm)/sqrt(n()))

penguin_f_summary

leveneTest(flipper_length_mm ~ species, data = penguins_f)
```
 
### Beeswarm plot with mean +- 1sd

```{r}
ggplot() + geom_beeswarm(data = penguins_f, 
                         aes(x= species, y = flipper_length_mm, 
                             color = "gray70"),
                             size = 3,
                             alpha = 0.5) +
  geom_point(data = penguin_f_summary, aes( x= species, y = mean_fl)) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax = mean_fl + sd_fl,
                    ymin = mean_fl - sd_fl),
                width = 0.1)
  
```

## One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguins_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```

Type III Anova with the car package: Use if you have unbalance design, uneven sample sizes.
```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguins_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```

